---
title: "Week2.Rmd"
author: "Sumit Kant"
date: "3 May 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Caret Package in R

## Caret package FUnctionality
* Preprocessing
* Data Splitting
* Testing/Training functions - train, predict 
* Model comparision - confusion matrix

## Machine learning Algorithms in R
* Linear Discriminant analysis
* Regression
* Naive Bayes
* Support Vector Machines
* Classfication and regression trees
* random forests
* Boosting

## Why Caret
Gives us a unified framework to make predictions just using one function despite having multiple packages of machine learninig algorithms

# EXAMPLE: SPAM Dataset

## Data Splitting
we have SPAM data set from the kernlab package. using the createDataPartition() function the data set is divided in testing and training set where 75% is training and rest is testing. The splitting variable is the type which contains whether the mail is market spam or not.

```{r warning=FALSE}
library(caret)
library(kernlab)
data(spam) # part of kernlab package
View(spam)
# Create data partition
inTrain <- createDataPartition(y = spam$type, p = 0.75, list = FALSE)
training <- spam[inTrain,]
testing <- spam[-inTrain,]
dim(training)
```
## Fitting MOdel
Training the data with general linear model
```{r warning=FALSE}
modelFit <- train(type~., data = training, method="glm")
modelFit
```

## Final Model
```{r warning=FALSE}
modelFit$finalModel
```
## Prediction
```{r warning=FALSE}
predictions <- predict(modelFit, newdata = testing)
predictions
```

## Confusion Matrix
```{r warning=FALSE}
confusionMatrix(predictions, testing$type)
```